---
title: "Exploratory Analysis"
output:
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(ggrepel)
library(hrbrthemes)
library(stringr)

theme_set(theme_ipsum_rc())

source("../code/lib.R")

knitr::opts_chunk$set(fig.width = 10, collapse = TRUE, echo = FALSE)
```
# MISSING MIGRANTS

```{r, include=FALSE}
missing_migrants = read_csv("../data/pretty/missing-migrants.csv")
```

The initial dataset contains `r NROW(missing_migrants)` data items and the following data attributes.

```{r results='asis'} 
descriptions = c("The region in which the incident took place",
                 "Estimated date of death",
                 "The year in which the incident occurred",
                 "The sum of the number of dead and missing",
                 "The number of females found dead or missing. If unknown, it is left blank.",
                 "The number of males found dead or missing. If unknown, it is left blank.",
                 "The number of individuals under the age of 18 found dead or missing. If unknown, it is left blank.",
                 "The number of migrants that survived the incident, if known",
                 "The determination of conditions resulting in the migrant's death",
                 "Place where the death(s) occurred or where the body or bodies were found",
                 "Place where the death(s) occurred or where the body or bodies were found",
                 "Articles that reported the incident")

cat(paste(paste0("- **", names(missing_migrants)),
          descriptions, sep = "** "), sep = '\n')
``` 

One of the dimensions I intend to investigate is the **information richness**. So I need a dataset with as much complete data as possible. The first decision is considering only the rows that contain the number of dead women and children. 

```{r}
missing_migrants = missing_migrants %>%
  filter(#!is.na(women), !is.na(children),
         year > 2016)
```

The resulting dataset contains `r NROW(missing_migrants)` now.

```{r}
missing_migrants %>%
  count(region) %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = reorder(region, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Number of incidents per region",
         x = NULL, y = "Frequency")
```

```{r}
migrants_sample = missing_migrants %>%
  filter(region == "Middle East") %>%
  arrange(desc(year)) %>%
  head(3)
```




```{r}
missing_migrants %>%
  mutate(year = as.factor(year)) %>%
  ggplot() +
  geom_jitter(aes(x = year, y = deaths), color = "red", width = 0.2) +
  geom_jitter(aes(x = year, y = survivors), color = "blue", width = 0.2) +
  scale_y_sqrt() + facet_wrap(~region) +
  labs(title = "Deads and survivors per region",
       subtitle = "Each incident corresponds to two dots: red for dead and blue for survivors",
       y = "Number of deads and survivors per incident",
       x = "Year")

ggsave("deads_and_survivors_per_region.png", width = 10, height = 8)
```

```{r}
data_sample = missing_migrants %>%
  filter(!is.element(region, c("Caribbean",
                               "Central America",
                               "Horn of Africa",
                               "North Africa",
                               "South America",
                               "Central Asia",
                               "East Asia",
                               "South Asia",
                               "North America",
                               "US-Mexico Border",
                               "Europe", # Removed to avoid bias (participants are European),
                               "Mediterranean" # Removed because it has much more data than the others
                               ))) %>%
  group_by(region, year) %>%
  summarise(deaths = sum(deaths, na.rm = T), survivors = sum(survivors, na.rm = T)) %>%
  mutate(death_survivor_rate = deaths / survivors)

data_sample %>%
  ggplot(aes(x = region, y = death_survivor_rate)) +
    geom_bar(stat = "identity")
```

```{r}
data_sample %>%
  ggplot(aes(x = year, y = death_survivor_rate, color = region)) +
    geom_line(size = 1.1) +
    geom_point() +
    geom_hline(yintercept = 0, size = 1.3, alpha = 0.5) +
    scale_x_continuous(breaks = seq(2016, 2018, 1))
```

```{r}
data_sample = data_sample %>%
  filter(!is.element(region, c("North Africa",
                               "Horn of Africa",
                               "US-Mexico Border",
                               "Sub-Saharan Africa"
                               )))
data_sample %>%
  ggplot(aes(x = year, y = death_survivor_rate, color = region)) +
    geom_line(size = 1.1) +
    geom_point() +
    geom_hline(yintercept = 0, size = 1.3, alpha = 0.5) +
    scale_x_continuous(breaks = seq(2016, 2018, 1))
```

```{r}
data_sample = missing_migrants %>%
  filter(year > 2015,
         is.element(region, c("Caribbean",
                              "Central America",
                              "Middle East",
                              "Southeast Asia")))
data_sample %>%  
  ggplot() +
    geom_jitter(aes(x = year, y = deaths), color = "red") +
    geom_jitter(aes(x = year, y = survivors), color = "blue") +
    scale_y_sqrt() + scale_x_continuous(breaks = seq(2016, 2018, 1)) +
    facet_wrap(~region)
```


# DEATHS AT THE BORDERS

## Identification of incidents

```{r}
dataset = import_dataset("../data/pretty/deaths-at-the-borders.csv")

# dataset %>%
#   count(incident_number) %>%
#   filter(n > 10, !is.na(incident_number)) %>%
#   ggplot(aes(x = reorder(incident_number, n), y = n)) +
#     geom_bar(stat = "identity") +
#     coord_flip() +
#     labs(title = "Most tragic incidents in the dataset", x = NULL, y = "Number of victims") +
#     theme(panel.grid.major.y = element_blank()) +
#     annotate("text", x = 28.5, y = 275, label = "atop(bold('Lampedusa boat tragedy'), 'There was a huge mobilization to find the victims')", parse = T) +
#     geom_segment(data = data.frame(x1 = 30.5, y1 = 275, x2 = 33.5, y2 = 275),
#                  aes(x = x1, y = y1, xend = x2, yend = y2), size = 1, color = "grey60")
```


```{r}
dataset %>%
  ggplot(aes(x = year, color = country)) +
    geom_line(stat = "count") +
    scale_y_sqrt() +
    labs(title = "Deaths at the borders during the years",
         x = "Year", y = "Number of victims", color = "Country")
```

```{r}
dataset %>%
  count(route) %>%
  ggplot(aes(x = reorder(route, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Deaths per route",
         x = "Route", y = "Number of victims")
```



```{r}
dataset %>%
  count(circumstances) %>%
  filter(!is.na(circumstances), n > 5) %>%
  ggplot(aes(x = reorder(circumstances, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Most frequent circumstances",
         x = NULL, y = "Frequency")
```


```{r}
word_count = data.frame(count_words(dataset$details_of_incident))
word_count$word <- rownames(word_count)
word_count %>%
  rename(n = 1) %>%
  filter(!grepl(pattern = "[0-9]+", x = word)) %>%
  slice(1:50) %>%
  ggplot(aes(x = 1, y = 1, size = n, label = word)) +
    geom_text_repel(segment.size = 0, force = 100) +
    scale_size(range = c(2, 10), guide = FALSE) +
    scale_y_continuous(breaks = NULL) +
    scale_x_continuous(breaks = NULL) +
    labs(x = '', y = '')
```


```{r}
dataset %>%
  count(nationality) %>%
  filter(!is.na(nationality), n > 5) %>%
  ggplot(aes(x = reorder(nationality, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Most frequent nationalities",
         x = NULL, y = "Frequency")
```


```{r}
dataset %>%
  count(descriptions_of_race) %>%
  filter(!is.na(descriptions_of_race)) %>%
  ggplot(aes(x = reorder(descriptions_of_race, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Most frequent descriptions of race",
         x = NULL, y = "Frequency")
```

```{r}
dataset %>%
  count(sex) %>%
  filter(!is.na(sex)) %>%
  ggplot(aes(x = reorder(sex, n), y = n)) +
    geom_bar(stat = "identity") +
    coord_flip() +
    labs(title = "Most frequent sex",
         x = NULL, y = "Frequency")
```

```{r}
dataset %>%
  filter(!is.na(age)) %>%
  ggplot(aes(x = age)) +
    geom_histogram() +
    labs(title = "Ages of victims",
         x = "Age", y = "Frequency")
```

